Measuring Perceptions and Preferences for Meritocracy

### 2021-06-03 paper_study3.Rmd
Estudio 3
Se ajusta preparación del S3, se crean 2 bases data_s3_conv y data_s3_inv, para convergencia e invarianza.

Se ajusta paper_s3, reproducible.

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather confusing. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization. Based on this critical analysis, we propose a survey instrument called *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis using data from an online survey study (N=2,141). Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

  |                                      | Social Origin  |     Hard Work    |     Wherewithal    |
  |--------------------------------------|----------------|------------------|--------------------|
  |     Meritocratic     Perception      |     -0.1975    |     0.1877       |     0.3921         |
  |     Unmeritocratic    Perception     |     0.3547     |     -0.0775      |     -0.1844        |
  |     Meritocratic    Preference       |     0.1078     |     0.0779       |     0.0819         |
  |     Unmeritocratic    Preference     |     -0.0397    |     0.0824       |     0.2082         |
  Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.

@@

```{r invargroup, eval=TRUE, echo=FALSE}
load(file = "../output/tables/tb.inv-group.RData")
col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")

kable(tab.inv, col.names = col.nam, align = "l",
      booktabs=TRUE,format = table_format,escape = FALSE,
      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
  footnote(general = footnote, footnote_as_chunk = T)
```

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
#bibliography:
#  - ../input/bib/EconomiaMoral.bib
#  - ../input/bib/meritocracy.bib
#  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

```{r val-conv}
pacman::p_load(kableExtra, polycor)

load(file="input/data/proc/data_s3_conv.RData")

pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
valores1 <- round(c(pl1,pl2,pl3,pl4),4)

pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)

pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)

variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")

table01 <- cbind(variables, valores1, valores2, valores3)

kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
    full_width = T,
    position = "center",
    font_size = 14,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", )

```

@@

# Analysis

```{r invargroup}

#---Load packages and data -------------------------------------#

pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,corrplot,psych,knitr,kableExtra,stats)

load(file = "input/data/proc/pov01_fact02.RData")

pov01 <- data_s3_inv %>% filter(dataset=="pvw01")
fs01  <- data_s3_inv %>% filter(dataset=="fsw02")

#-----Meassurement model--------------------------------------#

model01 <- '
perc_merit  =~ perc_effort + perc_talent
perc_nmerit =~ perc_wpart  + perc_netw
pref_merit  =~ pref_effort + pref_talent
pref_nmerit =~ pref_wpart  + pref_netw'

#-----Invariance---------------------------------------------#

inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE)

conf  <- inv01$fit.configural
weak  <- inv01$fit.loadings
strong<- inv01$fit.intercepts
strict<- inv01$fit.residuals

tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>% 
  as_tibble() %>% 
  select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>% 
  mutate(stars=gtools::stars.pval(pvalue),
         chisqt=paste0(round(Chisq,2)," (",Df,") "),
         decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
         model=c("Configural","Weak","Strong","Strict"))

fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])

# compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
fit.meas<- fit.meas %>% 
  mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
         diff.df   = df       - lag(df,   default = first(df)),
         diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
         diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
  round(3) %>% 
  mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))

tab.inv<- bind_cols(tab01,fit.meas) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>% 
  mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)

#clean values
tab.inv[tab.inv == c("0 (0) ")] <- NA
tab.inv[tab.inv == c(0)] <- NA

### Check an alternative method
configural        <- cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
weak.invariance   <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
strong.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
strict.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")

table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", **p < 0.001")

kable(tab.inv, col.names = col.nam, align = "l",
      booktabs=TRUE,format = table_format,escape = FALSE,
      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
  footnote(general = footnote, footnote_as_chunk = T)

```
